---
title: "Reproducible Research: Peer Assessment 1"
author: "oavasquez20"
date: "06-11-2020"
output: html_document
---

Extract Load and process data (ETL)

```{r ETL, message=FALSE,warning=FALSE}
#Libraries
library(ggplot2)
library(dplyr)

#Set Working directory
setwd("D:/Trabajo/R/Curso5/Project1/")

#Download file
Url <- "https://d396qusza40orc.cloudfront.net/repdata%2Fdata%2Factivity.zip"

#Im working on W10 so Method=Curl is not necesary :)
download.file(Url, destfile = paste0(getwd(), '/repdata_data_activity.zip'))

# Unzip file
unzip("repdata_data_activity.zip")
      
#Read
activity <- read.csv("activity.csv")

# Set date format
activity$date <- as.POSIXct(activity$date, "%Y%m%d")

# Get days
day <- weekdays(activity$date)

# Combine dataset and weekday
activity <- cbind(activity, day)
```

Summary data processed

```{r Summary,warning=FALSE}
summary(activity)
```

What is mean total number of steps taken per day?

```{r Question1,message=FALSE,warning=FALSE}
# Compute total steps taken per day
activity_TotalSteps <- with(activity, aggregate(steps, by = list(date), 
                                                sum, na.rm = TRUE))
# Col names
names(activity_TotalSteps) <- c("Date", "Steps")

# Plotting a histogram using ggplot2
g <- ggplot(activity_TotalSteps, aes(x = Steps)) + 
  geom_histogram(breaks = seq(0, 25000, by = 2500), fill = "#83CAFF", 
                 col = "black") + 
  ylim(0, 30) + 
  xlab("Total Steps Taken Per Day") + 
  ylab("Frequency") + 
  ggtitle("Total number of steps taken per Day")
```

Including Plots

```{r plot, echo=FALSE}
print(g)
```

Calculate and report the mean and median of the total number of steps taken per day

```{r mean}
mean(activity_TotalSteps$Steps)
```

```{r median}
median(activity_TotalSteps$Steps)
```

What is the average daily activity pattern?

```{r DailyActivity0,message=FALSE,warning=FALSE}
# Compute average number of steps taken, averaged across all days 
#by 5-min intervals.
DailyActivity_aggr <- aggregate(activity$steps, by = list(activity$interval), 
                                  FUN = mean, na.rm = TRUE)
# Col names
names(DailyActivity_aggr) <- c("Interval", "Mean")

# Plot on ggplot2
DailyActivity <- ggplot(DailyActivity_aggr, mapping = aes(Interval, Mean)) + 
  geom_line(col = "blue") +
  xlab("Interval") + 
  ylab("Average Number of Steps") + 
  ggtitle("Average Number of Steps Per Interval")
```

Plot Daily Activity

```{r DailyActivity, echo=FALSE}
print(DailyActivity)
```

Which 5-minute interval, on average across all the days in the dataset, contains the maximum number of steps?

```{r DailyActivity1}
DailyActivity_aggr[which.max(DailyActivity_aggr$Mean), ]$Interval
```

Imputing missing values

1. Calculate and report the total number of missing values in the 
dataset (i.e. the total number of rows with \color{red}{\verb|NA|}NAs)

```{r DailyActivity2}
sum(is.na(activity$steps))
```

2. Devise a strategy for filling in all of the missing values in the dataset. 
The strategy does not need to be sophisticated. For example, you could use 
the mean/median for that day, or the mean for that 5-minute interval, etc.

```{r DailyActivity3,message=FALSE,warning=FALSE}
imputedSteps <- DailyActivity_aggr$Mean[match(activity$interval, 
                                              DailyActivity_aggr$Interval)]
#Create a new dataset that is equal to the original dataset but 
#with the missing data filled in.

# Transfor steps in activity if they were missing values with the filled values from above.
activityImputed <- transform(activity, 
                             steps = ifelse(is.na(activity$steps), 
                                            yes = imputedSteps, 
                                            no = activity$steps))

# Create new dataset with imputed missing values.
totalActivityImputed <- aggregate(steps ~ date, activityImputed, sum)

# Col names
names(totalActivityImputed) <- c("date", "dailySteps")
```

4. Make a histogram of the total number of steps taken each day 
and Calculate and report the mean and median total number of 
steps taken per day. Do these values differ from the estimates 
from the first part of the assignment? What is the impact of 
imputing missing data on the estimates of the total daily number of steps?

```{r DailyActivity4, echo=FALSE}
# Plotting a histogram using ggplot2
p <- ggplot(totalActivityImputed, aes(x = dailySteps)) + 
  geom_histogram(breaks = seq(0, 25000, by = 2500), fill = "#83CAFF", 
                 col = "black") + 
  ylim(0, 30) + 
  xlab("Total Steps Taken Per Day") + 
  ylab("Frequency") + 
  ggtitle("Total Number of Steps Taken per Day")
print(p)
```

Mean of the total number of steps taken per day is:

```{r DailyActivityMean}
mean(totalActivityImputed$dailySteps)
```  

Median of the total number of steps taken per day is:

```{r DailyActivityMedian}
median(totalActivityImputed$dailySteps)
```  


Are there differences in activity patterns between weekdays and weekends?
For this part the \color{red}{\verb|weekdays()|}weekdays() function may 
be of some help here. Use the dataset with the filled-in missing values 
for this part.

1. Create a new factor variable in the dataset with two levels – “weekday” 
and “weekend” indicating whether a given date is a weekday or weekend day.

```{r Week,message=FALSE,warning=FALSE}
# Updating format of the dates
activity$date <- as.Date(strptime(activity$date, format="%Y-%m-%d"))

# Creating a function that distinguises weekdays from weekends
#My interface is in Spanish so I have to ask for "sábado" and "domingo"
#for weekends days

activity$dayType <- sapply(activity$date, function(x) {
  if(weekdays(x) == "sábado" | weekdays(x) == "domingo")
  {y <- "Weekend"}
  else {y <- "Weekday"}
  y
})

``` 

2. Make a panel plot containing a time series plot 
(i.e. \color{red}{\verb|type = "l"|}type = "l") of the 5-minute interval 
(x-axis) and the average number of steps taken, averaged across all weekday 
days or weekend days (y-axis). See the README file in the GitHub repository 
to see an example of what this plot should look like using simulated data.

```{r ByDay,message=FALSE,warning=FALSE}
# Creating the data set that will be plotted
activityByDay <-  aggregate(steps ~ interval + dayType, activity, mean, na.rm = TRUE)

# Plotting using ggplot2
dayPlot <-  ggplot(activityByDay, aes(x = interval , y = steps, color = dayType)) + 
  geom_line() + ggtitle("Average Daily Steps by Day Type") + 
  xlab("Interval") + 
  ylab("Average Number of Steps") +
  facet_wrap(~dayType, ncol = 1, nrow=2) +
  scale_color_discrete(name = "Day Type")
```

```{r Dayplot, echo=FALSE}
print(dayPlot) 
```













